How Kids Code and How We Know:\\An Exploratory Study on the Scratch Repository

### 2016-04-22 Paper/paper.tex
Changed lines of code to number of blocks

\begin{table*}[ht]
	\centering
	\begin{tabular}{lrrrrrr}
		\hline
		&\textbf{mean}&\textbf{min}&\textbf{Q1}&\textbf{median}&\textbf{Q3}&\textbf{max}\\
		\hline
		Sprites with code per project&5.68&1&1&2&5&525\\
		Scripts per project&17.35&1&2&5&12&3038\\
		Lines of code (LOC) per project&154.55&1&12&29&76&34622\\
		LOC in Stage per project&4.80&0&0&0&3&2613\\
		LOC in Sprites per project&115.57&0&10&26&68&34613\\
		LOC in Procedures per project&34.17&0&0&0&0&20552\\
		McCabe Cyclomatic Complexity (CC) per script&1.58&1&1&1&1&246\\
		McCabe CC per procedure script&3.75&1&1&2&4&183\\
		Procedures per project with procedures&11.50&1&1&2&6&847\\
		Arguments per Procedure&0.95&0&0&0&1&53\\
		Numerical arguments per procedure with arguments&1.73&0&1&1&2&22\\
		Text arguments per procedure with arguments&0.28&0&0&0&1&24\\
		Boolean arguments per procedure with arguments&0.13&0&0&0&0&14\\
		Calls per procedure&2.14&0&1&1&2&526\\
		Scripts with calls per procedure&1.13&0&1&1&1&59\\
		Variables per project&2.06&0&0&0&1&340\\
		Scripts utilizing variable&4.97&1&1&3&5&1127\\
		Lists per project&0.55&0&0&0&0&319\\
		Conditional statements per project&10.02&0&0&0&3&5950\\
		Recursive loop statements per project&7.65&0&1&2&5&2503\\
		User input blocks per project&4.77&0&0&1&4&1889\\
		Broadcast-receive statements per project&8.57&0&0&0&2&2460\\	
		\hline
	\end{tabular}
	\caption{Summary statistics from the dataset of \nScriptPrograms~non-empty Scratch projects}
	\label{tbl-size}
\end{table*}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/1sprites}
	\caption{Histogram of the number of sprites in the analyzed projects}
	\label{fig:size-sprites}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/1scripts}
	\caption{Histogram of the number of scripts in the analyzed projects}
	\label{fig:size-scripts}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/1locs}
	\caption{Histogram of the lines of code in the analyzed projects}
	\label{fig:size-locs}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/3cyclomatic}
	\caption{Histogram of the McCabe cyclomatic complexity of the \nscripts~analyzed scripts}
	\label{fig:cyclomatic}
\end{figure}

In Table \ref{tbl-size} we summarize statistics for the analyzed metrics. We use the mean value and the five-number summary to describe the dataset in terms of the number of sprites with code per project (including the stage sprite) and the number of scripts per project. We furthermore measure the lines of code per project, considering each Scratch block to represent exactly one line of code, thus not counting the closing constructs of loops or conditional statements as separate lines. Figures \ref{fig:size-sprites}, \ref{fig:size-scripts}, \ref{fig:size-locs} and plot the distribution of these size metrics.

We find that the majority of Scratch projects are small; 75\% of the projects have up to 5 sprites, 12 scripts and 76 lines of code, while one fourth of the projects have up to 12 lines of code. On the other end, 5\% of the projects (11,712) have more than 18 sprites and 4.8\% (11,214) consist of more than 500 lines of code. The analysis also highlighted some surprisingly large projects: 135 with more than 300 sprites and even 30 projects with more than 20,000 lines of code, whose Scratch identifiers are made available for further inspection.\footnotemark[\ref{repo}]

The lines of code metric was further analyzed to understand code organization. As shown in Table \ref{tbl-size}, the majority of Scratch code---74.78\% out of \nLOC~lines of code---is written within sprites. An additional 3.1\% of the total lines are found in the stage class. More interestingly, the remaining 22.11\% are lines within defined procedures, which are found in only 7.7\% (17,979) of the projects. The projects that contain procedures use them a lot; almost half of their total lines of code (48.81\%) are lines within procedures.

@@

\noindent
\fbox{
	\begin{minipage}{0.46\textwidth}
		RQ1: The majority of Scratch projects are small and simple; 75\% of the projects have up to 5 sprites, 12 scripts, 76 lines of code and no decision points. Most code is written in sprites. A small number of projects (7.7\%) use procedures, but they use them a lot and for more complex code. There exist surprisingly large and complex projects.
	\end{minipage}}

@@

\begin{figure}
	\begin{center}
		\includegraphics[width=6cm]{fig/Unconnected_clones.png}
		\caption{Cloned blocks that are not connected to an event. Program id: 12237615}
		\label{fig:Unconnected_clones}
	\end{center}
\end{figure}

>>>>

\begin{table*}[ht]
	\centering
	\begin{tabular}{lrrrrrr}
		\hline
		&\textbf{mean}&\textbf{min}&\textbf{Q1}&\textbf{median}&\textbf{Q3}&\textbf{max}\\
		\hline
		Sprites with code per project&5.68&1&1&2&5&525\\
		Scripts per project&17.35&1&2&5&12&3038\\
		Number of blocks per project&154.55&1&12&29&76&34622\\
		Blocks in Stage per project&4.80&0&0&0&3&2613\\
		Blocks in Sprites per project&115.57&0&10&26&68&34613\\
		Blocks in Procedures per project&34.17&0&0&0&0&20552\\
		McCabe Cyclomatic Complexity (CC) per script&1.58&1&1&1&1&246\\
		McCabe CC per procedure script&3.75&1&1&2&4&183\\
		Procedures per project with procedures&11.50&1&1&2&6&847\\
		Arguments per Procedure&0.95&0&0&0&1&53\\
		Numerical arguments per procedure with arguments&1.73&0&1&1&2&22\\
		Text arguments per procedure with arguments&0.28&0&0&0&1&24\\
		Boolean arguments per procedure with arguments&0.13&0&0&0&0&14\\
		Calls per procedure&2.14&0&1&1&2&526\\
		Scripts with calls per procedure&1.13&0&1&1&1&59\\
		Variables per project&2.06&0&0&0&1&340\\
		Scripts utilizing variable&4.97&1&1&3&5&1127\\
		Lists per project&0.55&0&0&0&0&319\\
		Conditional statements per project&10.02&0&0&0&3&5950\\
		Recursive loop statements per project&7.65&0&1&2&5&2503\\
		User input blocks per project&4.77&0&0&1&4&1889\\
		Broadcast-receive statements per project&8.57&0&0&0&2&2460\\	
		\hline
	\end{tabular}
	\caption{Summary statistics from the dataset of \nScriptPrograms~non-empty Scratch projects}
	\label{tbl-size}
\end{table*}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/1sprites}
	\caption{Histogram of the number of sprites in the analyzed projects}
	\label{fig:size-sprites}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/1scripts}
	\caption{Histogram of the number of scripts in the analyzed projects}
	\label{fig:size-scripts}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/1locs}
	\caption{Histogram of the number of blocks in the analyzed projects}
	\label{fig:size-locs}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{fig/charts/3cyclomatic}
	\caption{Histogram of the McCabe cyclomatic complexity of the \nscripts~analyzed scripts}
	\label{fig:cyclomatic}
\end{figure}

In Table \ref{tbl-size} we summarize statistics for the analyzed metrics. We use the mean value and the five-number summary to describe the dataset in terms of the number of sprites with code per project (including the stage sprite) and the number of scripts and blocks per project. Figures \ref{fig:size-sprites}, \ref{fig:size-scripts}, \ref{fig:size-locs} and plot the distribution of these size metrics.

We find that the majority of Scratch projects are small; 75\% of the projects have up to 5 sprites, 12 scripts and 76 blocks, while one fourth of the projects have up to 12 blocks. On the other end, 5\% of the projects (11,712) have more than 18 sprites and 4.8\% (11,214) consist of more than 500 blocks. The analysis also highlighted some surprisingly large projects: 135 with more than 300 sprites and even 30 projects with more than 20,000 blocks, whose Scratch identifiers are made available for further inspection.\footnotemark[\ref{repo}]

The number of blocks metric was further analyzed to understand code organization. As shown in Table \ref{tbl-size}, the majority of Scratch code---74.78\% out of \nLOC~blocks---is written within sprites. An additional 3.1\% of the total blocks are found in the stage class. More interestingly, the remaining 22.11\% are blocks within defined procedures, which are found in only 7.7\% (17,979) of the projects. The projects that contain procedures use them a lot; almost half of their total blocks (48.81\%) are within procedures.

@@

\noindent
\fbox{
	\begin{minipage}{0.46\textwidth}
		RQ1: The majority of Scratch projects are small and simple; 75\% of the projects have up to 5 sprites, 12 scripts, 76 blocks and no decision points. Most code is written in sprites. A small number of projects (7.7\%) use procedures, but they use them a lot and for more complex code. There exist surprisingly large and complex projects.
	\end{minipage}}

@@

\begin{figure}
	\begin{center}
		\includegraphics[width=6cm]{fig/Unconnected_clones.png}
		\caption{Cloned blocks that are not connected to an event. Project id: 12237615}
		\label{fig:Unconnected_clones}
	\end{center}
\end{figure}
